---
title: "Robust test 2"
author: "Zhong"
date: "2025-06-05"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
#This is a data set consisting of measurements of water salinity (i.e., its salt concentration) and river discharge taken in North Carolina's Pamlico Sound, recording some bi-weekly averages in March, April, and May from 1972 to 1977. This data is available in the robustbase package, data(salinity)

#(a):Model this relationship using:OLS and GMestimator KS2014setting ,Tabulate the estimated coefficients and their standard errors for both fits.

#1.load the data
```{r}
## Load the robustbase package
library(robustbase)
#input data
data("salinity")
#view structure of data 
head(salinity)
str(salinity)
#step 2 using ols fit model
#OLS fit
ols_fit <- lm(Y~X1+X2+X3,data = salinity)
#summary OLS fit
summary(ols_fit)
#step 3 using  GMestimator KS2014setting
gm_fit <- lmrob(Y~X1+X2+X3,data = salinity,setting="KS2014")
#"KS2014"是一套预设的鲁棒控制参数，来自 Koller & Stahel (2014)，适用于很多实际问题，鲁棒性好，自动配置好一整套鲁棒回归参数
summary(gm_fit)
#step 4 tabulate coefficients and standard errrors

results <- data.frame(
  ols_coef = coef(ols_fit),#data.frame 里面是=，不是<-，注意区别！！！
  gm_coef = coef(gm_fit),
  ols_se = summary(ols_fit)$coefficients[,2],
  gm_se = summary(gm_fit)$coefficients[,2]#这里不要加逗号！！！
)
print(results)


```

#residual polts of  two models 
```{r}
# 1. 设置一个图形窗口，画两张图并排
par(mfrow = c(1, 2))

# 2. OLS 残差图
plot(ols_fit$fitted.values, resid(ols_fit),#plot散点图，横轴拟合值，纵轴残差
     main = "OLS Residuals",#图表名称
     xlab = "Fitted values",#X轴名称
     ylab = "Residuals",#Y轴名称
     col = "red", pch = 19)#点的形状是实心圆（代码 19 是一个点样式）
abline(h = 0, lty = 2)#abline在图上画一条直线，h=0表示y = 0 的水平线，ity=2表示这条线是虚线（线型类型 2）

# 3. GM 残差图
plot(gm_fit$fitted.values, resid(gm_fit),
     main = "GM (KS2014) Residuals",
     xlab = "Fitted values",
     ylab = "Residuals",
     col = "blue", pch = 19)
abline(h = 0, lty = 2)

```



#（b）Identify outliers and classify observations into vertical, good and bad leverage points.

```{r}
#step 1 计算残差和杠杆值，cook distance
library(car)
df <- salinity
df$restudent <- rstudent(ols_fit)#残差（即实际值与拟合值的差）非常大（标准化残差 > 2 或 < -2）

df$hat <- hatvalues(ols_fit)# 杠杆值 hᵢ
df$cook <- cooks.distance(ols_fit)# Cook's D，用于衡量单个点对拟合结果的影响力

#step 2 设定判断标准
n <- nrow(salinity)       # 总观测数
p <- length(coef(ols_fit))    # 参数数（包括截距）
lev.thresh <- 2 * p / n   # 杠杆值阈值
inf.thresh <- 4 / n       # Cook's D 阈值

#step 3 找出异常值
outliers <- subset(df,abs(restudent)>2)#找出外部学生化残差的绝对值 > 2 的点，这些是可能的离群点。

leverage_1 <-subset(df,hat>lev.thresh)#找出杠杆值大于阈值的点，即高杠杆点
influential_1 <- subset(df,cook>inf.thresh)#找出Cook's 距离大于阈值的点，这些是可能会显著改变拟合结果的有影响力观测值。

#step 4 可视化表格化这些极端点

print(outliers)
print(leverage_1)
print(influential_1)

```

#(c)(i) Compare parameter estimates, standard errors, and the set of flagged outliers.
## 汇总异常值编号
```{r}
print("Coefficient Estimates and Standard Errors:")
coefs <- summary(ols_fit)$coefficients
print(coefs)
```


#(ii) Explain why the GM weight function reduces the influence of the bad leverage observations you identified in part (b) and hence shifts the fitted surface.

#Answer:It down-weights large residuals If an observation doesn’t fit the model (i.e., it lies far from the fitted surface), it gets a small weight.

#It further down-weights high-leverage points with large residuals
#A point far in X-space and far off the surface is seen as highly suspect. GM estimators give it very little influence in estimating the regression surface.

#This reduces their "pulling" power
#In OLS, all points are treated equally (full weight = 1). In GM, bad leverage points might get a weight close to 0. As a result, they don’t distort the regression fit.

#The GM weight function reduces the influence of bad leverage points by assigning them small weights. Bad leverage observations have both high leverage (extreme X values) and large residuals. In GM estimation, such points receive low weights because their residuals are large, and sometimes because their leverage is high. As a result, they contribute very little to the estimation of the regression surface. This prevents them from pulling the fitted surface toward themselves, shifting the surface toward the main data cluster and improving robustness.